Flowchart of Current Algorithm

### 2009-06-21 docs/flowchart/flowchart.tex
pass in gravity to make_w0_spherical
fix the lower boundary condition index for w0 in spherical
update paper IV info

Here we outline the algorithm currently implemented in the code.  Our
starting point for this description are the series of papers describing
the development of the algorithm:
\begin{itemize}
\item {\em Low Mach Number Modeling of Type Ia
  Supernovae. I. Hydrodynamics,} A.~S.~Almgren, J.~B.~Bell, 
  C.~A.~Rendleman, \& M.~Zingale 2006, ApJ, 637, 922 (henceforth
  paper~I)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
  Evolution,} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
  2006, ApJ, 649, 927 (henceforth paper~II)\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
  2008, ApJ, 684, 449 (henceforth paper~III)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale,
  in preparation (henceforth paper~IV)
\end{itemize}
We also carry over some ideas from our small-scale low Mach number algorithm
for astrophysical flows:
\begin{itemize}
\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
J.~B.~Bell, M.~S.~Day, C.~A.~Rendleman, S.~E.~Woosley, \& M.~Zingale
2004, JCP, 195, 2, 677 (henceforth BDRWZ)
\end{itemize}

@@

\item {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:

@@

\noindent We discretize this elliptic equation in the radial dimension as:
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j} -
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j-1} \right\}
+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{j-\myhalf} \nonumber\\
&=& 
- \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
   - \left( \frac{4 \ow \rho_0 g}{r} \right)_{j-\myhalf}
   - \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip ,
\end{eqnarray}
where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
Then, using hydrostatic equilibrium, we expand this to 
\begin{eqnarray}
& \frac{1}{\Delta r} \left\{
 \left( \frac{\gammabar p_0}{r^2}\right)_{j  } \frac{\left[ (r^2 \dw)_{j+\myhalf} - (r^2 \dw)_{j-\myhalf} \right]}{\Delta r} -\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \frac{\left[ (r^2 \dw)_{j-\myhalf} - (r^2 \dw)_{j-\thalf}\right]}{\Delta r} \right\} & \nonumber \\
&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf} = \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf} & \nonumber \\
& - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
  \left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip, &
\end{eqnarray}
where $\etarho$ at cell-centers is the average of the neighboring edge-values.  
If we write this in matrix form, so that:
\begin{equation}
A_j (r^2 \dw)_{j-\thalf} + B_j (r^2 \dw)_{j-\myhalf} + C_j (r^2 \dw)_{j+\myhalf} = F_j\enskip,
\end{equation}
then:
\begin{eqnarray}
A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1}  \enskip, \\
B_j &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j  }  
               +\left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1} \right] 
              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) \enskip , \\
C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j}  \enskip, \\
F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
 - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
\left[ \left( r^2 \etarho^{\inp} \right)_{j} - \left( r^2 \etarho^{\inp}  \right)_{j-1} \right] 
\nonumber \\
&& - \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip .
\end{eqnarray}
Here,
\begin{equation}
g_{j-\myhalf} = \frac{G}{r_{j-\myhalf}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\myhalf}^3 - r_{k-\thalf}^3) \; \rho_{0,k-1}^{\inp} \enskip.
\end{equation}

We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
corresponds to $j=1,$ by setting:
\begin{eqnarray}
A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
B_1 &=& 1\enskip.
\end{eqnarray}
We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $j=N,$
by setting:
\begin{eqnarray}
A_N &=& -1\enskip, \nonumber \\
B_N &=&  1\enskip, \nonumber \\
C_N = F_N &=& 0\enskip.
\end{eqnarray}
Finally, $w_0^{\outp} = \ow + \delta w_0$.
\end{description}

@@

If we are using the volume discrepancy term, compute 
$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
\item For spherical geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
\rhozero^n,p_0^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
\end{description}

@@

\item Define:
\begin{equation}
\gammabar^{\nph} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip , ~~~
\rhozero^{\nph} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip , ~~~
p_0^{\nph} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip ,
\end{equation}
with
\begin{equation}
 \gammabar^{n+1,\star} = {\rm{\bf Avg}} 
\left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, X_k^{n+1,\star}\right) \right] 
\enskip .
\end{equation}

\item Compute $w_0^{n+\myhalf}$:
\begin{description}
\item For planar geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph}},
\gammabar^{n+\myhalf},p_0^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.

>>>>

Here we outline the algorithm currently implemented in the code.  Our
starting point for this description are the series of papers describing
the development of the algorithm:
\begin{itemize}
\item {\em Low Mach Number Modeling of Type Ia
  Supernovae. I. Hydrodynamics,} A.~S.~Almgren, J.~B.~Bell, 
  C.~A.~Rendleman, \& M.~Zingale 2006, ApJ, 637, 922 (henceforth
  paper~I)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
  Evolution,} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
  2006, ApJ, 649, 927 (henceforth paper~II)\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
  2008, ApJ, 684, 449 (henceforth paper~III)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
M.~Zingale, A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& S.~E.~Woosley
  2009, ApJ, submitted (henceforth paper~IV)
\end{itemize}
We also carry over some ideas from our small-scale low Mach number algorithm
for astrophysical flows:
\begin{itemize}
\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
J.~B.~Bell, M.~S.~Day, C.~A.~Rendleman, S.~E.~Woosley, \& M.~Zingale
2004, JCP, 195, 2, 677 (henceforth BDRWZ)
\end{itemize}

@@

\item {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},g^\inp,\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:

@@

\noindent We discretize this elliptic equation in the radial dimension as:
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j} -
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j-1} \right\}
+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{j-\myhalf} \nonumber\\
&=& 
- \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
   - \left( \frac{4 \ow \rho_0 g}{r} \right)_{j-\myhalf}
   - \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip ,
\end{eqnarray}
where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
Then, using hydrostatic equilibrium, we expand this to 
\begin{eqnarray}
& \frac{1}{\Delta r} \left\{
 \left( \frac{\gammabar p_0}{r^2}\right)_{j  } \frac{\left[ (r^2 \dw)_{j+\myhalf} - (r^2 \dw)_{j-\myhalf} \right]}{\Delta r} -\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \frac{\left[ (r^2 \dw)_{j-\myhalf} - (r^2 \dw)_{j-\thalf}\right]}{\Delta r} \right\} & \nonumber \\
&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf} = \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf} & \nonumber \\
& - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
  \left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip, &
\end{eqnarray}
where $\etarho$ at cell-centers is the average of the neighboring edge-values.  
If we write this in matrix form, so that:
\begin{equation}
A_j (r^2 \dw)_{j-\thalf} + B_j (r^2 \dw)_{j-\myhalf} + C_j (r^2 \dw)_{j+\myhalf} = F_j\enskip,
\end{equation}
then:
\begin{eqnarray}
A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1}  \enskip, \\
B_j &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j  }  
               +\left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1} \right] 
              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) \enskip , \\
C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j}  \enskip, \\
F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
 - \frac{g_{j-\myhalf}^\inp}{r_{j-\myhalf}^2 \Delta r} 
\left[ \left( r^2 \etarho^{\inp} \right)_{j} - \left( r^2 \etarho^{\inp}  \right)_{j-1} \right] 
\nonumber \\
&& - \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip .
\end{eqnarray}
We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
corresponds to $j=0,$ by setting:
\begin{eqnarray}
A_0 = C_0 = F_0 &=& 0\enskip, \nonumber \\
B_0 &=& 1\enskip.
\end{eqnarray}
We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $j=N,$
by setting:
\begin{eqnarray}
A_N &=& -1\enskip, \nonumber \\
B_N &=&  1\enskip, \nonumber \\
C_N = F_N &=& 0\enskip.
\end{eqnarray}
Finally, $w_0^{\outp} = \ow + \delta w_0$.
\end{description}

@@

If we are using the volume discrepancy term, compute 
$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
\item For spherical geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
\rhozero^n,p_0^n,g^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
\end{description}

@@

\item Define:
\begin{eqnarray}
\gammabar^{\nph} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip, \qquad
\rhozero^{\nph} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip, \nonumber \\
p_0^{\nph} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip, \qquad
g^\nph = \frac{g^n + g^{n+1,\star}}{2}\enskip ,
\end{eqnarray}
with
\begin{equation}
 \gammabar^{n+1,\star} = {\rm{\bf Avg}} 
\left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, X_k^{n+1,\star}\right) \right] 
\enskip .
\end{equation}

\item Compute $w_0^{n+\myhalf}$:
\begin{description}
\item For planar geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph}},
\gammabar^{n+\myhalf},p_0^{n+\myhalf},g^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.
